---
title: "R Notebook for hypothesis tests"
output:
  html_document:
    toc: yes
    toc_float: TRUE
    theme: united
    df_print: paged
---
Set up hypotheses first. Then, only keep the P-values for the correct hypthesis test 
(left sided, right sided, two-sided test). Make sure that the sample evidence
supports the altenative hypothesis; otherwise the null hypothesis is accepted
by default.


# One sample means test with with std dev unknown (t-test)
```{r}
#Input Data
hypoMean<-0
sampleN<-10
xbar<-0.084
sampleStdDev<-0.0844
#Computations
  #Std Dev of xbar (std error, SE)
xbarSE<-sampleStdDev/sqrt(sampleN)
Tstat<-(xbar-hypoMean)/xbarSE
#print the t statistic
Tstat  
# Choose the right P-value according to the null hypothesis
# P-value for left-tailed test
#Pvalue<-pt(Tstat,sampleN-1)
# P-value for right-tailed test
Pvalue<-1-pt(Tstat,sampleN-1)
# P-value for two-tailed test
#Pvalue<-2*pt(-abs(Tstat),sampleN-1)
Pvalue
```



# One sample proportion test (Z-test)
```{r}
#Input Data
hypoProp<-0.25
sampleN<- nrow(myCD)
sampleProp<-sum(myCD$Manufacturer == "Kellogg")/nrow(myCD)
#Computations
 #Compute std dev of sample proportion (Std error of sample proportion)
myPropSE<-sqrt(sampleProp*(1-sampleProp)/sampleN)
Zstat<-(sampleProp-hypoProp)/myPropSE
#Print the Z statistic
Zstat
# Choose the right P-value according to the null hypothesis
#P-value for left-tailed test
#Pvalue<-pnorm(Zstat)
#P-value for right-tailed test
Pvalue<-(1-pnorm(Zstat))
#P-value for two-tailed test
#Pvalue<-2*pnorm(-abs(Zstat))
#Print the P-value
Pvalue
```



# Two sample means test 
This two sample measn test assumes "equal" variances (Pooled t-test)
```{r}
#Input Data
hypoDiff<-0.0
sampleN1<-40
sampleN2<-37
xbar1<-5.3
xbar2<-6.5
#sampleSD1<-1.99
#sampleSD2<-1.16
sampleV1<-sqrt(1.99)
sampleV2<-sqrt(2.36)
#Computations with Standard Daviation
#pooledVar<-((sampleN1-1)*sampleSD1^2 + (sampleN2-1)*sampleSD2^2)/(sampleN1+sampleN2-2)
#Tstat<-(xbar1-xbar2-hypoDiff)/sqrt(pooledVar*(1/sampleN1+1/sampleN2))

#Computations with Variance
pooledVar<-((sampleN1-1)*sampleV1^2 + (sampleN2-1)*sampleV2^2)/(sampleN1+sampleN2-2)
Tstat<-(xbar1-xbar2-hypoDiff)/sqrt(pooledVar*(1/sampleN1+1/sampleN2))
#print the t statistic
Tstat  
# Choose the right P-value according to the null hypothesis
# P-value for left-tailed test
#Pvalue<-pt(Tstat,sampleN1+sampleN2-2)
# P-value for right-tailed test
#Pvalue<-1-pt(Tstat,sampleN1+sampleN2-2)
# P-value for two-tailed test
Pvalue<-2*pt(-abs(Tstat),sampleN1+sampleN2-2)
Pvalue
```

# Paired t test for differences.

Compute the difference for each pair of observations. Compute sample mean and sample standard deviation for the differences. Then apply the one sample t test on the differences (see first script).



# Two sample proportions test (Z-test)
```{r}
#Input Data
hypoDiff<-0.0
sampleP1<-63/78
sampleP2<-49/82
pooledP<-(63+49)/(78+82)
sampleN1<-78
sampleN2<-82
#Computations
Zstat<-(sampleP1-sampleP2-hypoDiff)/sqrt(pooledP*(1-pooledP)*(1/sampleN1+1/sampleN2))
#Print the Z statistic
Zstat
# Choose the right P-value according to the null hypothesis
#P-value for left-tailed test
#Pvalue<-pnorm(Zstat)
#P-value for right-tailed test
#Pvalue<-(1-pnorm(Zstat))
#P-value for two-tailed test
Pvalue<- 2 * pnorm(-abs(Zstat))
#Print the P-value
Pvalue
```